\documentclass[12pt]{article}

\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{marvosym}
\usepackage{graphicx}
\usepackage{fontawesome}
\usepackage{xcolor}

\pagestyle{empty}

\usepackage[scale=0.775]{geometry}

\setlength{\parindent}{0pt}
\addtolength{\parskip}{6pt}

\def\firstname{Xavier}
\def\familyname{Loizeau}
\def\FileAuthor{\firstname \familyname}
\def\FileTitle{\firstname \familyname's cover letter for Zenob\"e}
\def\FileSubject{Cover letter}
\def\FileKeyWords{\firstname \familyname, Cover letter}

\renewcommand{\ttdefault}{pcr}

\usepackage{url}
\urlstyle{tt}
\ifpdf
  \usepackage[pdftex,pdfborder=0,breaklinks,baseurl=http://,pdfpagemode=None,pdfstartview=XYZ,pdfstartpage=1]{hyperref}
  \hypersetup{
    pdfauthor   = \FileAuthor,%
    pdftitle    = \FileTitle,%
    pdfsubject  = \FileSubject,%
    pdfkeywords = \FileKeyWords,%
    pdfcreator  = \LaTeX,%
    pdfproducer = \LaTeX}
\else
  \usepackage[dvips]{hyperref}
\fi


\begin{document}
\sffamily   % for use with a résumé using sans serif fonts;
%\rmfamily  % for use with a résumé using serif fonts;
\hfill%
\begin{minipage}[t]{.6\textwidth}
\raggedleft%
{\bfseries Xavier Loizeau, \\ PhD, MIMA}\\[.35ex]
\small\itshape%
Flat1, 16 Acre road,\\
KT2 6EF, Kingston upon Thames,\\
England\\[.35ex]
\faMobile~ (+44) 7845 079595\\
\faAt~\href{mailto:subscription@loizeau.eu}{subscription@loizeau.eu}
\end{minipage}\\[1em]
%
\begin{minipage}[t]{.4\textwidth}
\raggedright%
{\bfseries Zenob\"e}\\
\end{minipage}
\hfill % US style
%\\[1em] % UK style
\begin{minipage}[t]{.4\textwidth}
\raggedleft % US style
\today
%April 6, 2006 % US informal style
%05/04/2006 % UK formal style
\end{minipage}\\
\raggedright

To whom this concerns,

I am delighted to hereby apply to join Zenob\"e.
I am particularly interested in your offer of Senior EV fleet analyst, but also open to other opportunities you would offer me.
After working for nearly five years at the National Physical Laboratory, I intend to take the next important step in my career.
I am looking for the opportunity to take on more responsibilities in leadership and strategy.
While doing so, I want to remain an active scientist participating in exciting, scientifically challenging projects at the service of society and to be part of a dynamic, inclusive team where all members feel encouraged to give their best while being themselves.

\medskip

Zenob\"e's mission impacts two essential contemporary problems I am passionate about climate change and the accessibility of our cities.
Building a healthier environment today by reducing greenhouse gas emissions and mitigating the effects of climate change by providing resilient ways of life is an essential motivator in my career.
Contributing to the elaboration of tools that empower citizens in fighting climate change, such as you do, is something I would be honoured to participate in.
In addition, I am particularly pleased to see how closely you collaborate with public transport companies, which represent the most viable future means of transport in our cities.
In particular, environmentally friendly means of transport should be accessible regardless of income, and public transports are the fastest way to this goal.

\medskip

My experience in multi-disciplinary project teams and diverse application areas is a key skill that I would use in your company.
I use state-of-the-art data science techniques with various data types and communicate their results to diverse audiences.
An essential ingredient to my successful collaborations is my pride in delivering rigorous, reproducible and reusable work by applying quality processes.
I also contribute to the strategic vision of the laboratory by designing project proposals, applying for internal and external grants, being an active member of special interest groups, involving myself in the design of internal policies, leading recruitment interviews, and designing training material for recruits.
For all those reasons, I am confident I can positively contribute to your organisation and would be delighted to discuss our potential collaboration further with you soon.

\medskip

Respectfully yours,

Xavier Loizeau

\vfill%
%{\slshape Enclosure}
%{\slshape Attachment: curriculum vit\ae{}}
\end{document}
