\documentclass[12pt]{article}

\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{marvosym}
\usepackage{graphicx}
\usepackage{fontawesome}
\usepackage{xcolor}

\pagestyle{empty}

\usepackage[scale=0.775]{geometry}

\setlength{\parindent}{0pt}
\addtolength{\parskip}{6pt}

\def\firstname{Xavier}
\def\familyname{Loizeau}
\def\FileAuthor{\firstname \familyname}
\def\FileTitle{\firstname \familyname's cover letter for SMC Insights – The Strategy Workshop}
\def\FileSubject{Cover letter}
\def\FileKeyWords{\firstname \familyname, Cover letter}

\renewcommand{\ttdefault}{pcr}

\usepackage{url}
\urlstyle{tt}
\ifpdf
  \usepackage[pdftex,pdfborder=0,breaklinks,baseurl=http://,pdfpagemode=None,pdfstartview=XYZ,pdfstartpage=1]{hyperref}
  \hypersetup{
    pdfauthor   = \FileAuthor,%
    pdftitle    = \FileTitle,%
    pdfsubject  = \FileSubject,%
    pdfkeywords = \FileKeyWords,%
    pdfcreator  = \LaTeX,%
    pdfproducer = \LaTeX}
\else
  \usepackage[dvips]{hyperref}
\fi


\begin{document}
\sffamily   % for use with a résumé using sans serif fonts;
%\rmfamily  % for use with a résumé using serif fonts;
\hfill%
\begin{minipage}[t]{.6\textwidth}
\raggedleft%
{\bfseries Xavier Loizeau, \\ PhD, MIMA}\\[.35ex]
\small\itshape%
Flat1, 16 Acre road,\\
KT2 6EF, Kingston upon Thames,\\
England\\[.35ex]
\faMobile~ (+44) 7845 079595\\
\faAt~\href{mailto:subscription@loizeau.eu}{subscription@loizeau.eu}
\end{minipage}\\[1em]
%
\begin{minipage}[t]{.4\textwidth}
\raggedright%
{\bfseries The Alan Turing Institute}\\
\end{minipage}
\hfill % US style
%\\[1em] % UK style
\begin{minipage}[t]{.4\textwidth}
\raggedleft % US style
\today
%April 6, 2006 % US informal style
%05/04/2006 % UK formal style
\end{minipage}\\
\raggedright

To whom this concerns,

I hereby apply with great enthusiasm to join the Alan Turing Institute.
After working for nearly five years at the National Physical Laboratory (NPL), I intend to take the next important step in my career.
I am looking for the opportunity to take on more responsibilities in leadership and strategy.
While doing so, I want to remain an active researcher participating in exciting, scientifically challenging projects and to be part of a dynamic, inclusive team where all members feel encouraged to give their best while being themselves.

\medskip

The Turing is an ideal place to keep on growing.
Your grand challenges, "Transformation of healthcare" and "Environment and sustainability", particularly resonate with me.
Both topics are at the forefront of my concerns and fields of research I participate in with great interest.
The capabilities you identify as core, such as "foundational capability in data science and AI" and "open source infrastructure", are essential in delivering a trustworthy AI that serves our entire society.
Hence, I am confident that the ways of working at The Turing would suit me well.
In addition, your "Equality Diversity and Inclusion" page shows an inclusive organisation, which is essential to letting all express their true potential.

\medskip

During the last four years, I have developed my network and area of speciality (metrology for complex objects) by contributing to varied scientific projects.
I adapted to different application areas such as climate science, pathology, and mass spectrometry imaging, which led me to use many state-of-the-art data science techniques.
I take pride in delivering rigorous, reproducible and reusable work by applying quality processes based on tools such as Git, GitLab and CI/CD pipelines, and I develop automation scripts and templates for the rest of the laboratory to encourage the adoption of these tools.
I also contribute to the strategic vision of the laboratory by designing project proposals, applying for internal and external grants, being an active member of special interest groups, involving myself in the design of internal policies, leading recruitment interviews, and designing training material for recruits.
For all those reasons, I am confident I can positively impact your organisation and would be delighted to discuss our potential collaboration further with you soon.

\medskip

Respectfully yours,

Xavier Loizeau

\vfill%
%{\slshape Enclosure}
%{\slshape Attachment: curriculum vit\ae{}}
\end{document}